Universals of word order result from optimization of grammars for efficient communication

### 2019-02-25 second-draft/4-paper-names.tex
revising

In this work, we show that the word order of languages has evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.

@@

\section{DLM}

\begin{figure*}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/four-grammars-dlm.png}
	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure. In this example, Grammar 1 achieves the lowest overall dependency length. Indeed, this is the }
	\label{fig:dlm-grammars}
\end{figure*}

The sentences in Figure~\ref{fig:arabic-japanese} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words. For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} if the verb, rather than e.g. the subject or an indirect object. A dependency arc is drawn from a \key{head} (e.g. \emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g. \emph{tegami-o}). Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}. Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.

In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels. This is because, according to the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.

@@

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{figures/grammars.png}
	\caption{Given a set of dependency structures, grammars define a set of sentences. For simplicity, we use English words. Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb. \textbf{TODO This figure is just a draft sketch so far.}}
	\label{fig:grammars}
\end{figure}

Dependency Length Minimization is heuristic and stipulative.

@@

We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
In Shannon's model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
In Figure~\ref{fig:comm}, the message consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.

@@

All natural languages have some degree of word order regularities that are specified in their grammar.
For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
That is, speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
%Speakers encode dependency trees into sentences according to a grammar~(Figure~\ref{fig:grammars}).
%While natural languages differ in the freedom of word order, all languages rely on some amount of word order reglarity to encode meaning.
This is illustrated in Figure~\ref{fig:grammars}: Grammars specify how dependency structures are encoded into strings of words.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.
%In a speech community, speakers and listeners have shared knowledge of these rules, which allows the listener to decode the dependency tree.
%In the terminology of information theory, these rules specify a \emph{channel} through which the speaker transmits the dependency structures.

@@

%For answering these questions,

@@

We now show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.

Confirming previous theoretical arguments, we see that optimizing for dependency length correctly accounts for all word order correlations. 
Predictability and parseability predict four and six correlations, respectively, making complementary predictions.
Efficiency correctly predicts all the word order correlations.

@@

>>>>

In this work, we show that the word orders of languages have evolved to optimize efficiency, and that efficiency optimization accounts for the Greenbergian word order correlations.

@@

\section{Dependency Length Minimization}

Some prior work has has suggested a functional explanation of Greenberg's correlations in terms of Dependency Length Minimization \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
On the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, it has been confirmed that languages reduce dependency length.
This has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.

(CITE) have made theoretical arguments that dependency length minimization should produce word order correlations.
Here, we present the first large-scale test of this idea.

\begin{figure}
\centering

\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
	   I \&	   visited \& friends  \\
   \end{deptext}
	%   \deproot{3}{ROOT}
   \depedge{2}{1}{nsubj}
   \depedge{2}{3}{obj}
\end{dependency}

\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
	   I \&	   went   \& to \& school  \\
   \end{deptext}
	%   \deproot{3}{ROOT}
   \depedge{2}{1}{nsubj}
   \depedge{2}{3}{}
   \depedge{3}{4}{}
   %\depedge[arc angle=50]{7}{6}{ATT}
\end{dependency}

\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
	   I \&	   wrote \& letters   \& to \& friends  \\
   \end{deptext}
	%   \deproot{3}{ROOT}
   \depedge{2}{1}{nsubj}
   \depedge{2}{3}{obj}
	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
   \depedge{2}{4}{}
   \depedge{4}{5}{}
   %\depedge[arc angle=50]{7}{6}{ATT}
\end{dependency}

	\caption{Dependency trees}\label{fig:dep-trees}
\end{figure}

The sentences in Figure~\ref{fig:dep-trees} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words. For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} of the verb, rather than e.g. the subject or an indirect object. A dependency arc is drawn from a \key{head} (e.g. \emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g. \emph{tegami-o}). Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}. Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.

The dependency length of a sentence is the sum of the lengths of its syntactic dependencies \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
\begin{equation}
    \label{eq:depl}
\operatorname{DepL}(t; {\bf w}) = \sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
\end{equation}
where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost.

All natural languages have some degree of word order regularities that are specified in their grammar.
For instance, English places subjects before and objects after the verb (Figure~\ref{fig:dep-trees}).
Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
This is illustrated in Figure~\ref{fig:dlm-grammars}: Grammars specify how dependency structures are encoded into strings of words.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.

We create optimized grammars that minimize the average dependency length.

%See SI for our method for creating optimized grammars.

\begin{figure*}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/four-grammars-dlm.png}
	\caption{The DLM theory of word order universals: We show how four possible grammars can order the given dependency structure. In this example, Grammar 1 achieves the lowest overall dependency length. Indeed, this is the Given a set of dependency structures, grammars define a set of sentences. For simplicity, we use English words. Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb. \textbf{TODO This figure is just a draft sketch so far.}}
	\label{fig:dlm-grammars}
\end{figure*}

%In order to interpret a sentence correctly, a listener must be able to reconstruct the information expressed in the dependency tree and its labels. This is because, according to the principle of compositionality \cite{frege1892sinn}, the meaning of a sentence is a function of the meanings of the parts and how they are combined; the dependency tree (including the labels on the arcs) specifies how the meanings of words are combined.

@@

\paragraph{Results}
Now present the correlations and the results.
We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.
Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.

Confirming previous theoretical arguments, we see that optimizing for dependency length correctly accounts for all word order correlations.

@@

We have verified that dependency length minimization produces Greenberg's correlation universals.
However, dependency length is both heuristic and stipulative.
The actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}. 
It is also stipulative: There is no a priori reason why dependency length should be the crucial determining factor determining word order.
There are other models of processing cost that make different predictions.

We show that the word order predictions of DLM arise from deeper principles of efficient communication.

@@

%\begin{figure}
%    \centering
%    \includegraphics[width=0.45\textwidth]{figures/grammars.png}
%	\caption{Given a set of dependency structures, grammars define a set of sentences. For simplicity, we use English words. Each grammar defines a consistent ordering for the different kinds of syntactic relations, e.g., Grammars 1 and 2 order the object after the verb, Grammars 3 and 4 order the object before the verb. \textbf{TODO This figure is just a draft sketch so far.}}
%	\label{fig:grammars}
%\end{figure}
%
%
%

@@

We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
In Shannon's model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
These are encoded in the dependency structure.
%Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
%This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
In Figure~\ref{fig:comm}, the message consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.

@@

Speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences
For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.

@@

We use the same 51 corpora.

@@

We now show that efficiency optimization predicts Greenberg's word order correlations. % in optimized grammars. 
Predictability and parseability predict four and six correlations, respectively, making complementary predictions.
Efficiency correctly predicts all the word order correlations.

@@

\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
	   katabt \& risāla \& li \& sadīq  \\
	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
	   wrote \& letter   \& to \& friend  \\
   \end{deptext}
	%   \deproot{3}{ROOT}
   \depedge{1}{2}{obj}
	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
   \depedge{1}{3}{}
   \depedge{3}{4}{}
   %\depedge[arc angle=50]{7}{6}{ATT}
\end{dependency}
\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
      tomodachi \& ni \& tegami-o \& kaita \\
 \textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
	   friend \& to \& letter  \& wrote \\
   \end{deptext}
	%   \deproot{3}{ROOT}
   \depedge{4}{2}{}
	%   \depedge[edge start x offset=-6pt]{2}{5}{ATT}
   \depedge{2}{1}{}
   \depedge{4}{3}{obj}
   %\depedge[arc angle=50]{7}{6}{ATT}
\end{dependency}
